A Hybrid Approach for News Recommender System using Optimization Methods

### 2021-03-11 paper/paper.tex
updated paper

\title{A hybrid approach for news recommender system using optimization methods}

@@

@@

@@

        \begin{itemize}
            \item \Cref{sec:related} describes the relevant related work
            \item \Cref{sec:input} describes input data
            \item \Cref{sec:overview} explains our modular design and architecture
            \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
            \item \Cref{sec:evaluation} provides tests and experiments validating
our system’s results
            \item \Cref{sec:further} explains future work
            \item \Cref{sec:summary} presents conclusions
        \end{itemize}

@@

\begin{itemize}
    \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users. It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
    \item \textbf{Content-based filtering}: is based on the assumption that people who liked items with certain attributes in the past, will like the same kind of items in the future as well. It makes use of item features to compare the item with user profiles and provide recommendations
    \item \textbf{Session filtering}: is selecting cadidates based on user's activity on current session
    \item \textbf{Popularity filtering}: uses information such as number of views, shows, comments, etc.
    \item \textbf{Demographic filtering}: uses demographic data such as age, gender, education, etc. for identifying categories of users
    \item \textbf{Time-based filtering}: ranking news is a way that more recent items have higher scores
\end{itemize}

@@

\subsection{Collaborative filtering}

There were many studies on this topic, but paper \citep{evaluation} proves that well-tuned basic SVD++ approach beats newely presented algorithms.

@@

As we solving domain specific task, we have domain specific data.

@@

@@

        1 & 2021-01-08 22:08:39 & 9  & politics, conflicts\\
        2 & 2021-01-09 10:28:58 & 5  & IT, social media\\
        3 & 2021-01-09 14:20:34 & 12 & accident\\
        \vdots & \vdots & \vdots & \vdots \\\bottomrule

@@

\begin{itemize}
    \item \emph{source\_id}: source of news item
\end{itemize}

@@

@@

    1  &  Azerbaijan denies reports on construction of Turkish air bases in the country & Information that Turkey will create air bases ...        \\
    2  &  Durov announced the massive transition of WhatsApp users to Telegram          & Telegram developer Pavel Durov said in his channel ...   \\
    3  &  Passenger plane that disappeared from radar crashed                           & Passenger plane taking off from Jakarta, disappeared ... \\
    \vdots & \vdots & \vdots \\\bottomrule

     \hline
    \end{tabular}
    }

    \caption{news item}
    \label{tab:item}
\end{table}

@@

@@

\begin{itemize}
    \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
    \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
\end{itemize}

@@

There is an option to react on item via leaving emotion and/or writing a comment.

@@

\begin{itemize}
    \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
\end{itemize}

@@

If \emph{user\_id} subscribed to the \emph{source\_id}.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{cc}
        \toprule
        \textit{word} & \textit{category} \\
        \midrule
        alexey\_nalalny & politics \\
        crash & accident \\
        telegeram & IT \\
        \vdots & \vdots \\
        \bottomrule
    \end{tabular}%
    \caption{words' categories}
    \label{tab:words_categories}
\end{table}

@@

For collaborative filtering recommendation we should have something known as user-item matrix which may be formed from user activity from tables (cite tables)

@@

@@

\section{Implementation}
\label{sec:implementation}

@@

    \begin{table}[]
        \centering
        \begin{tabular}{ccc}
            \toprule
            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
            \midrule
            1                 & 0.25                                  & 1                               \\
            2                 & 1                               & 0.5                                \\
            \vdots & \vdots & \vdots \\
            \bottomrule
            \end{tabular}%
        \caption{boost values}
        \label{tab:boost_values}
    \end{table}

@@

\subsection{Online}

\subsection{Offline}

>>>>

\title{A Hybrid Approach for News Recommender System using Optimization Methods}

@@

    Recommender systems have been studied to present items, such as movies, music, and books \citep{movies} \citep{films} \citep{books}. The term now has a broader connotation, describing any system that produces individualized recommendations as output or has the effect of guiding the user in a personalized way to interesting or useful objects in a large space of possible options. Such systems have an obvious appeal in an environment where the amount of on-line information vastly outstrips any individual’s capability to survey it.

    Recommender systems are now an integral part of some e-commerce sites such as Amazon.com and CDNow \citep{commerce}. It is the criteria of ’individualized’ and ’interesting and useful’ that separate the recommender system from information retrieval systems or search engines \citep{coin}. The semantics of a search engine are ’matching’: the system is supposed to return all those items that match the query ranked by degree of match.  Techniques such as relevance feedback enable a search engine to refine its representation of the user’s query, and represent a Simple form of recommendation.

     There are three main types of recommendation methods: memorybased, model-based, and hybrid \citep{survey}. Memory-based methods \citep{memory} usually use similarity metrics to obtain the distance between two users or two items. Model-based methods use demographic, content, or aggregated information to create a model that generates recommendations. Hybrid \citep{combining} methods combine different types of recommenders to gain better performance.

@@

    So we are going to present a hybrid recommender system.

@@

        \begin{itemize}
            \item \Cref{sec:related} describes the relevant related work
            \item \Cref{sec:input} describes input data
            \item \Cref{sec:overview} explains our modular design and architecture
            % \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
            \item \Cref{sec:evaluation} provides tests and experiments validating
our system’s results
            \item \Cref{sec:further} explains future work
            \item \Cref{sec:summary} presents conclusions
        \end{itemize}

@@

\begin{itemize}
    \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users. It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person; there were many studies on this topic, but paper \citep{evaluation} proves that well-tuned basic SVD++ approach beats newely presented algorithms
    \item \textbf{Content-based filtering}: is based on the assumption that people who liked items with certain attributes in the past, will like the same kind of items in the future as well. It makes use of item features to compare the item with user profiles and provide recommendations
    \item \textbf{Session filtering}: is selecting cadidates based on user's activity on current session
    \item \textbf{Popularity filtering}: uses information such as number of views, shows, comments, etc.
    \item \textbf{Demographic filtering}: uses demographic data such as age, gender, education, etc. for identifying categories of users
    \item \textbf{Time-based filtering}: ranking news is a way that more recent items have higher scores
\end{itemize}

Also there is several ways \citep{survey1} to combine recommenders between each other:

\begin{itemize}
    \item \textbf{Weighted}: The scores (or votes) of several recommendation techniques are.combined together to produce a single recommendation
    \item \textbf{Switching}: The system switches between recommendation techniques depending on the current situation
    \item \textbf{Mixed}: Recommendations from several different recommenders are presented at the same time
    \item \textbf{Feature combination}: Features from different recommendation data sources are thrown together into a single recommendation algorithm
    \item \textbf{Cascade}: Features from different recommendation data sources are thrown together into a single recommendation algorithm 
    \item \textbf{Feature augmentation}: Output from one technique is used as an input feature to another
    \item \textbf{Meta-level}: The model learned by one recommender is used as input to another
\end{itemize}

@@

@@

As we solving domain specific task, we have domain specific data.
Out data consist of information about news, so it includes following tables:

@@

The metadata about news. Here we have an \textit{item\_id} column, which stands for unique item indentifier, afterwards \textit{date}, that shows when this item was released. The \textit{source\_id} column stands for id of a publisher of this news item. \textit{category} column means category of current news item. This category is taken from news text by keywords.

@@

        1 & 2021-01-08 22:08:39 & 9  & politics, conflicts\\
        2 & 2021-01-09 10:28:58 & 5  & IT, social media\\
        3 & 2021-01-09 14:20:34 & 12 & accident\\
        \vdots & \vdots & \vdots & \vdots \\\bottomrule
    \end{tabular}
    % }

@@

% \begin{itemize}
%     \item \emph{source\_id}: source of news item
% \end{itemize}

@@

Content table hold news texts in a \textit{news\_content} column. This is the primary information that is being used by our recommender system, because we are able to extract a lot of valuable data from text, such as topics.

@@

    1  &  Azerbaijan denies reports on construction of Turkish air bases in the country & Information that Turkey will create air bases ...        \\
    2  &  Durov announced the massive transition of WhatsApp users to Telegram          & Telegram developer Pavel Durov said in his channel ...   \\
    3  &  Passenger plane that disappeared from radar crashed                           & Passenger plane taking off from Jakarta, disappeared ... \\
    \vdots & \vdots & \vdots \\\bottomrule
    \end{tabular}
    }
    \caption{news item}
    \label{tab:item}
\end{table}

@@

As we operate with user activity we should have user logs, such as what items were clilcked at what time, so we have separate table that contains this information.    \emph{shows} stands for \emph{item\_id} was shown to the \emph{user\_id}, \emph{views} is if \emph{item\_id} was clicked by the \emph{user\_id}.

@@

% \begin{itemize}
%     \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
%     \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
% \end{itemize}

@@

We have an explicit feedback that user may leave on a news if he wants to. These options includes leaving emoji, which is one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\} or comment, which we can analyze afterwards.

@@

% \begin{itemize}
%     \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
% \end{itemize}

@@

Each piece of news is being posted by some feed and user have an ability to subscribe to the feed. So it may give us useful information if user prefers content from one feed to content to another feed. So following table illustrates if \emph{user\_id} subscribed to the \emph{source\_id}.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{cc}
        \toprule
        \textit{word} & \textit{category} \\
        \midrule
        trump & politics \\
        crash & accident \\
        telegeram & IT \\
        \vdots & \vdots \\
        \bottomrule
    \end{tabular}%
    \caption{words' categories}
    \label{tab:words_categories}
\end{table}

@@

For collaborative filtering recommendation we should have something known as user-item matrix which may be formed from user activity from tables (cite tables).

We use the modification of Funk MF, which factorized the user-item rating matrix as the product of two lower dimensional matrices, the first one has a row for each user, while the second has a column for each item. The row or column associated to a specific user or item is referred to as latent factors.

$r_{u i}=\sum_{f=0}^{n } H_{u, f} W_{f, i}$

While Funk MF is able to provide very good recommendation quality, its ability to use only explicit numerical ratings as user-items interactions constitutes a limitation. Modern day recommender systems should exploit all available interactions both explicit (e.g. numerical ratings) and implicit (e.g. likes, purchases, skipped, bookmarked). To this end SVD++ was designed to take into account implicit interactions as well.  Compared to Funk MF, SVD++ takes also into account user and item bias.
The predicted rating user $u$ will give to item $i$ is computed as:

$r_{u i}=\mu+b_{i}+b_{u}+\sum_{f=0}^{n} H_{u, f} W_{f, i}$

@@

So when we apply information about news item popularity, we are able to give them scores via following algorithm:

\texttt{min-max} normalizing each of \emph{shows\_num}, \emph{views\_num}, \emph{emotions\_num}, \emph{comments\_num}, then dividing by 4 (to have 1 as max after sum), and sum all of these values.

latent Dirichlet allocation (LDA)

@@

same as previous

% \section{Implementation}
% \label{sec:implementation}

@@

    \begin{table}[h]
        \centering
        \begin{tabular}{ccc}
            \toprule
            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
            \midrule
            1                 & 0.25                                  & 1                               \\
            2                 & 1                               & 0.5                                \\
            \vdots & \vdots & \vdots \\
            \bottomrule
            \end{tabular}%
        \caption{boost values}
        \label{tab:boost_values}
    \end{table}

@@

% \subsection{Online}

% \subsection{Offline}
